---
output:
  html_document:
    fig_caption: yes
    keep_md: yes
---
---
title: "week_6_assignment"
author: "Adejare Windokun"
date: "Wednesday, October 01, 2014"
output: html_document


Will first of all load the required libraries, rvest and (devtools, if not already installed)

```{r}
if (!require(devtools)) install.packages('devtools')
library(devtools)

if (!require(rvest)) install_github("hadley/rvest")
library(rvest)

```

I will use Ebay (ebay.com) and a particular page on Ebay that sells the 9500ix radar detector:

http://www.ebay.com/sch/i.html?_from=R40&_trksid=p2050601.m570.l1313.TR11.TRC1.A0.H0.X9500ix&_nkw=9500ix&_sacat=0

The items of interest are:

Item of interest    CSS 
Name of the item    .lvtitle
Price               .lvprice or .g-b
Format of bids      .lvformat
Details             .lvdetails
Extraxs             .lvextras
Shipping            .lvshipping


```{r comment=""}
radar <- html("http://www.ebay.com/sch/i.html?_from=R40&_trksid=p2050601.m570.l1313.TR11.TRC1.A0.H0.X9500ix&_nkw=9500ix&_sacat=0")


radar_name <- radar %>% 
  html_nodes(".lvtitle") %>%
  html_text() %>%
  gsub("[\t\n\r\f\v]", "", .) 
radar_name

radar_price <- radar %>% 
  html_nodes(".lvprice") %>%
  html_text() %>%
  gsub("[\t\n\r\f\v]", "", .) %>%
  gsub("(^ +)|( +$)", "", .)
radar_price

radar_bid <- radar %>% 
  html_nodes(".lvformat") %>%
  html_text() %>%
  gsub("[\t\n\r\f\v]", "", .) 
radar_bid

radar_extra <- radar %>% 
  html_nodes(".lvextras") %>%
  html_text() %>%
  gsub(c("[\t\n\r\f\v]", "^\\s+|\\s+$"), "", .) 
radar_extra

radar_shipping <- radar %>% 
  html_nodes(".lvshipping") %>%
  html_text() %>%
  gsub("[\t\n\r\f\v]", '', .) %>%
  gsub("^\\s+|\\s+$", "", .)
radar_shipping


```

It is not possible to convert the results into a data frame because the length of the vectors differ. The way Ebay stores the price, there is a buy it now price and the current bid price and both of them get scraped leading to a difference in the 
length of the vectors. I haven't been able to figure out how to fix this.

Therefore to create a data frame I am going to use the same Ebay site but now limit the results to "Buy it now", which only has a single buy it now price

Website: http://www.ebay.com/sch/i.html?_from=R40&_sacat=0&_ipg=50&_nkw=9500ix&rt=nc&LH_BIN=1


```{r comment=""}
radar = html("http://www.ebay.com/sch/i.html?_from=R40&_sacat=0&_ipg=50&_nkw=9500ix&rt=nc&LH_BIN=1")

radar_name <- radar %>% 
  html_nodes(".lvtitle") %>%
  html_text() %>%
  gsub("[\t\n\r\f\v]", "", .) 
radar_name

radar_price <- radar %>% 
  html_nodes(".lvprice") %>%
  html_text() %>%
  gsub("[\t\n\r\f\v]", "", .) 
radar_price

radar_bid <- radar %>% 
  html_nodes(".lvformat") %>%
  html_text() %>%
  gsub("[\t\n\r\f\v]", "", .) 
radar_bid

radar_extra <- radar %>% 
  html_nodes(".lvextras") %>%
  html_text() %>%
  gsub("[\t\n\r\f\v]", "", .) %>%
  gsub("(^ +)|( +$)", "", .)
radar_extra

length(radar_name)
length(radar_price)
length(radar_bid)
length(radar_extra)

radar_df = data.frame(radar_name, radar_price, radar_bid, radar_extra)
radar_df

```

