---
title: "week_9_quiz"
author: "Adejare Windokun"
date: "Friday, October 24, 2014"
output: word_document
---


```{r}
if (!require(rmongodb)) install.packages('rmongodb')
library(rmongodb)
if (!require(jsonlite)) install.packages('jsonlite')
library(jsonlite)

# install rmongodbHelper package from GitHub
# install.packages("devtools")
if (!require(devtools)) install.packages('devtools')
library(devtools)

if (!require(rmongodbHelper)) devtools::install_github("joyofdata/rmongodbHelper")
library(rmongodbHelper)

# Example of the rmongodbHelper
# json_qry <-
# '{
#   "$or": [
#     {"a":1},
#     {"a":3}
#   ]
# }'
#  
# bson <- rmongodbHelper::json_to_bson(json_qry)

```

connect to the mongodb database
```{r}
mongo = mongo.create()
mongo.is.connected(mongo)

```

Code to destroy the database and collection if neccessary

```{r}
if (mongo.is.connected(mongo) == TRUE) {
     mongo.drop(mongo, "employment.employees")
     mongo.drop.database(mongo, "employment")
#     #res <- mongo.get.database.collections(mongo, "employment")
#     #print(res)
#     #close connection
#     mongo.destroy(mongo)
 }

```



1. Create a new MongoDB database called employment.

```{r}
db <- "employment"
mongo <- mongo.create(db=db)

```



2. Insert a new record for Wendy Yasquez into the database and into a collection called employees.

```{r}
ins = '{"Name":"Wendy Yasquez", "Title":"Assistant Professor", "Salary":86000, "Department":"Computer Science", "Hire_Year":1998}'

#This shows you what the JSON statement looks like before you insert it into the database
cat(prettify(ins))


if (mongo.is.connected(mongo)) {
        mongo.insert(mongo, "employment.employees", ins)
}

```

See what records you have


```{r}
#No query criteria
q = "{}"
if (mongo.is.connected(mongo)) {
    cursor <- mongo.find(mongo, "employment.employees",query = q)
    while (mongo.cursor.next(cursor))
        print(mongo.cursor.value(cursor))
    mongo.cursor.destroy(cursor)
}

```


3. Write a JavaScript function to insert new professors into the employees collection.
Could not do this in R - used Robomongo

db.system.js.save(
                   { _id: "insertProf",
                     value : function (name, title, salary, department, hire_year){ db.employees.insert({Name:name, Title:title, Salary:salary, Department:department, Hire_Year:hire_year})}
                   }
                 )


4. Use this function to insert the records for Raoul Dewan, Isabelle Winters, and Jack McDunn.
    First you have to load the scripts
    db.loadServerScripts();
    
    Did this using the function I created:
    
    insertProf('Raoul Dewan', 'Assistant Professor', 78000, ['Physics', 'Biology'], 2009) 
    insertProf('Isabelle Winters', 'Associate Professor', 92000, 'Physics', 1995)
    insertProf('Jack MuDunn', 'Associate Professor', 101000, 'Physics', 1993)
    
```{r}
ins1 = '{"Name":"Raoul Dewan", "Title":"Assistant Professor", "Salary":78000, "Department": "[Physics, Biology]", "Hire_Year":2009}'

ins2 = '{"Name":"Isabelle Winters", "Title":"Associate Professor", "Salary":92000, "Department":"Physics", "Hire_Year":1995}'

ins3 = '{"Name":"Jack MuDunn", "Title":"Associate Professor", "Salary":101000, "Department":"Physics", "Hire_Year":1993}'


        mongo.insert(mongo, "employment.employees", ins1)
        mongo.insert(mongo, "employment.employees", ins2)
        mongo.insert(mongo, "employment.employees", ins3)


```

    
5. Write a JavaScript function to insert new administrative employees into the employees collection.

db.system.js.save(
                  { _id: "insertAdmins",
                    value : function (name, title, salary, division, location, hire_year){ db.employees.insert({Name:name, Title:title, Salary:salary,
                    Division:division, Location:location, Hire_Year:hire_year})
                  }}
                )

 6. Use this function to insert the records for Tonja Baldner and Dennis Bohnet.
    db.loadServerScripts();
    insertAdmins('Tonja Baldner', 'Assistant to the Dean', 42000, 'Ats and Sciences', '', 2001)
    insertAdmins('Dennis Bohnet', 'Vice President', 106000, 'Academic Affairs', 'Main Campus', 1997)


```{r}
ins4 = '{"Name":"Tonja Baldner", "Title":"Assistant to the Dean", "Salary":42000, "Division":"Ats and Sciences", "Location": "", "Hire_Year":1995}'

ins5 = '{"Name":"Dennis Bohnet", "Title":"Vice President", "Salary":106000, "Division":"Academic Affairs", "Location": "Main Campus",  "Hire_Year":1993}'


        mongo.insert(mongo, "employment.employees", ins4)
        mongo.insert(mongo, "employment.employees", ins5)
        

```

    
7. Show the code that will return all employees with salaries less than $90,000.

db.employees.find({Salary:{$lt:90000}})

```{r}

q = '{"Salary":{"$lt":90000}}'
if (mongo.is.connected(mongo)) {
    cursor <- mongo.find(mongo, "employment.employees",query = q)
    while (mongo.cursor.next(cursor))
        print(mongo.cursor.value(cursor))
    mongo.cursor.destroy(cursor)
}

```


8. Show the code that will return all professors with salaries less than $90,000.

db.employees.find({Title: { $regex: /PROFESSOR/i }, Salary:{$lt:90000}})


9. Show the code that will return all Physics professors hired before 2001.

db.employees.find({Department:'Physics', Hire_Year:{$lt:2001}})


```{r}
json_qry <-
 '{
   "Department": "Physics",
     "Hire_Year":
        {"$lt":2001}
     
 }'

cat(prettify(json_qry))
bson <- rmongodbHelper::json_to_bson(json_qry)

if (mongo.is.connected(mongo)) {
    cursor <- mongo.find(mongo, "employment.employees", query = bson)
    while (mongo.cursor.next(cursor))
        print(mongo.cursor.value(cursor))
    mongo.cursor.destroy(cursor)
}

```


10. Show the code that will return all professors who teach for departments other than Physics. (This should include professors who teach for Physics and also other departments.)

db.employees.find({ $and: [
    {$or: 
        [
            {Department:{$ne:'Physics'}},
       
            {'Department.1': {$exists: true}}
       
        ]},
    
     {Title: { $regex: /PROFESSOR/i }}  

   ]} 
)

11. Show the code that will return all employees who were either hired before 1997 or who have salaries greater than $100,000.

db.employees.find({$or: [ {Salary:{$gt:100000}}, {Hire_Year:{$lt:1997}} ] } )


```{r}
json_qry <-
 '{
    "$or": [{
        "Salary": 
            {"$gt":100000}},
        {"Hire_Year":
            {"$lt":1997}} ]

 }'

cat(prettify(json_qry))
bson <- rmongodbHelper::json_to_bson(json_qry)

if (mongo.is.connected(mongo)) {
    cursor <- mongo.find(mongo, "employment.employees", query = bson)
    while (mongo.cursor.next(cursor))
        print(mongo.cursor.value(cursor))
    mongo.cursor.destroy(cursor)
}

```


12. Suppose Tonja Baldner has been given a 10% raise. Show the code that will update her salary correctly.
db.employees.update(
   { Name: 'Tonja Baldner'},
   { $mul: { Salary: 1.1 } }
)

```{r}
cursor <- mongo.find.one(mongo, "employment.employees", query = '{"Name":"Tonja Baldner"}')
print("Before Update")
print(cursor)

crit = '{ "Name": "Tonja Baldner"}'
obj = '{ "$mul": { "Salary": 1.1 } }'
    
mongo.update(mongo, "employment.employees", criteria = crit, objNew= obj)
cursor <- mongo.find.one(mongo, "employment.employees", query = '{"Name":"Tonja Baldner"}')
print("After Update")
print(cursor)

```

Have to do question 14 first, otherwise, Prof Dewan is already deleted from the database
14. Instead of removing Professor Dewan’s record, we might prefer to create a new collection called pastemployees and move his record there. Show the code that will move his record to the new collection and add a departyear value of 2014 to his record. (You can do it in two steps.)

Will first copy over the document from the employees collection to the newly created pastemployees collection
Will then insert the new field = departyear and update it to 2014

Do everything in one step:

db.employees.find({Name: 'Raoul Dewan'}).forEach(function(doc){
   db.pastemployees.insert(doc)
   db.pastemployees.update(doc,{$set : {"departyear":2014}})
   db.employees.remove(doc)
   
});



```{r}
cursor <- mongo.find.one(mongo, "employment.employees",query = '{"Name":"Raoul Dewan"}')

print (cursor)

mongo.insert(mongo, "employment.pastemployees", cursor)
crit = '{"Name":"Raoul Dewan"}'
objNew = '{"$set" : {"departyear":2014}}'
mongo.update(mongo, "employment.pastemployees", criteria = crit, objNew)

```

13. Professor Dewan has been offered a job at another university. Show the code that would remove his record from the database.
db.employees.remove( { Name : "Raoul Dewan" }, 1 )


```{r}
crit = '{ "Name": "Raoul Dewan"}'
mongo.remove(mongo, "employment.employees", criteria = crit)

```





