
---
title: "week_9_assignment"
author: "Adejare Windokun"
date: "Saturday, October 25, 2014"
output: word_document
---


Part 1: mongoimport
Code to import files

C:\Users\jare\SkyDrive\WorkDocs\CUNY\607\Week 9\mongodb\bin>mongoimport --host localhost --db unitedstates --collection states --type csv --headerline --file statedata.csv
connected to: localhost
2014-10-25T12:10:49.253-0400 check 9 51
2014-10-25T12:10:49.263-0400 imported 50 objects

C:\Users\jare\SkyDrive\WorkDocs\CUNY\607\Week 9\mongodb\bin>mongoimport --host localhost --db unitedstates --collection states --type tsv --headerline --file districtdata.txt
connected to: localhost
2014-10-25T12:11:09.303-0400 imported 1 objects

C:\Users\jare\SkyDrive\WorkDocs\CUNY\607\Week 9\mongodb\bin>mongoimport --host localhost --db unitedstates --collection states --type tsv --headerline --file inhabitedterritorydata.txt
connected to: localhost
2014-10-25T12:11:18.595-0400 imported 5 objects


```{r}
if (!require(rmongodb)) install.packages('rmongodb')
library(rmongodb)
if (!require(jsonlite)) install.packages('jsonlite')
library(jsonlite)
if (!require(plyr)) install.packages('plyr')
library(plyr)
# install rmongodbHelper package from GitHub
# install.packages("devtools")
if (!require(devtools)) install.packages('devtools')
library(devtools)

if (!require(rmongodbHelper)) devtools::install_github("joyofdata/rmongodbHelper")
library(rmongodbHelper)

# Example of the rmongodbHelper
# json_qry <-
# '{
#   "$or": [
#     {"a":1},
#     {"a":3}
#   ]
# }'
#  
# bson <- rmongodbhelper::json_to_bson(json_qry)

```

connect to the mongodb database
```{r}
mongo = mongo.create()
mongo.is.connected(mongo)

```

see all the records
```{r}
#No query criteria
q = "{}"
if (mongo.is.connected(mongo)) {
    cursor <- mongo.find(mongo, "unitedstates.states", query = q, limit = 10L)
    while (mongo.cursor.next(cursor))
        print(mongo.cursor.value(cursor))
    mongo.cursor.destroy(cursor)
}

```

Part 2: rmongodb
Query to select out the states { state: { $exists: true } }

```{r}
 
json_qry <-
 '{
   "state": 
     {"$exists":true}
 }'


bson <- rmongodbHelper::json_to_bson(json_qry)

states = data.frame()
cursor <- mongo.find(mongo, "unitedstates.states", bson)
i = 1
while(mongo.cursor.next(cursor))
{
    tmp = mongo.bson.to.list(mongo.cursor.value(cursor))
    tmp.df = as.data.frame(t(unlist(tmp)), stringAsFactors = FALSE)
    tmp.df = subset(tmp.df, select = c(-1))
    states = rbind.fill(states, tmp.df)
i = i + 1
}
head(states)
```

For the districts
Query to select out the districts { federal_district: { $exists: true } } 

```{r}

json_qry <-
 '{
   "federal_district": 
     {"$exists":true}
 }'


bson <- rmongodbHelper::json_to_bson(json_qry)


districts = data.frame()
cursor <- mongo.find(mongo, "unitedstates.states", bson)
i = 1
while(mongo.cursor.next(cursor))
{
    tmp = mongo.bson.to.list(mongo.cursor.value(cursor))
    tmp.df = as.data.frame(t(unlist(tmp)), stringAsFactors = FALSE)
    tmp.df = subset(tmp.df, select = c(-1))
    districts = rbind.fill(districts, tmp.df)
i = i + 1
}
head(districts)

```

For territories
Query to select out the territory { territory: { $exists: true } } 

```{r}



json_qry <-
 '{
   "territory": 
     {"$exists":true}
 }'


bson <- rmongodbHelper::json_to_bson(json_qry)

territories = data.frame()
cursor <- mongo.find(mongo, "unitedstates.states", bson)
i = 1
while(mongo.cursor.next(cursor))
{
    tmp = mongo.bson.to.list(mongo.cursor.value(cursor))
    tmp.df = as.data.frame(t(unlist(tmp)), stringAsFactors = FALSE)
    tmp.df = subset(tmp.df, select = c(-1))
    territories = rbind.fill(territories, tmp.df)
i = i + 1
}
head(territories)

```




